Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds

### 2021-01-17 Manuscript_Publication.Rmd
manuscript changes

---
title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds
author:
bibliography: References/SS_Extinction.bib
output:
  pdf_document:
    keep_tex: yes
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
  \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
  \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{}
  \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
  \raggedbottom
csl: References/evolution.csl
nocite: | 
  @Kraaijeveld_2011, @Maia_2013, @Huang_2014, @Gomes_2016, @Cooney_2017, @Janicke_2018, @Mason_2017, @Iglesias-Carrasco_2019, @Hosner_2020, @Price-Waldman_2020
subparagraph: yes
abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates. However, the net effect of sexual selection on diversification is hard to predict, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes. Theoretical work also suggests that the net effect of sexual selection on diversification should depend strongly on ecological factors, though this prediction has seldom been tested. Here, we test whether variation in sexual selection can predict speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera) and whether this relationship is mediated by environmental factors. Male-biased sexual selection, and specifically sexual size dimorphism, predicted two of the three measures of speciation rates that we examined. The link we observed between sexual selection and speciation was independent of environmental variability, though species with smaller ranges had higher speciation rates. There was no association between any proxies of sexual selection and extinction rate. Our findings support the view that male-biased sexual selection, as measured by frequent predictors of male-male competition, has shaped diversification in the largest radiation of birds.
---
<!-- \maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia} -->

@@

Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data [@Jetz_2012], together with significant advances in phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification. Furthermore, the diversity of outcomes and approaches in previous studies suggests that the association between species diversity and sexual selection is far from clear (reviewed in @tsuji2020sexual). **See Table 1 for a review of the major findings from previous studies testing the association between sexual selection and speciation**.

@@

Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014] because their biology and phylogenetic relationships are comparatively well-known. A 2011 meta-analysis covering 20 primary studies of birds and other taxa found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects or fish [@Kraaijeveld_2011]. However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers. More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].

Here, we investigate the association between sexual selection and diversification in birds while building upon previous approaches in multiple ways. We use two measures of the strength of sexual selection: sexual dichromatism [@Dale_2015], as well as an index of male-biased sexual selection [@Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care. We use these two measures because sexual dichromatism does not always signal the presence of strong sexual selection and *vice versa* [@Dale_2015]. For example, male and female dunnocks (*Prunella modularis*) are similarly coloured yet sexual selection appears to be strong [@davies1986re]. Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018]. Additionally, our analysis includes multiple ecological and environmental variables, allowing us to control for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].

@@

Because sexual dichromatism is presumably imperfectly correlated with variation in the strength of sexual selection across taxa, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the 'index of male-biased sexual selection'. This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care). The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015]. This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).  \phantomsection

@@

Diversification is the result of two processes, speciation and extinction through time. To estimate speciation rates, we first obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny. Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES; also referred to as diversification rate/DR) is derived from the sum of edge lengths from each tip towards the root, with each edge towards the root having the length down-weighted [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018]. Crucially, studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018]. Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path. These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE); but, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].

@@

We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds; \autoref{Speciation_Figure}). To do this, we tested the association between speciation/extinction and sexual selection, environmental variability and their interaction using phylogenetic least squares (PGLS) models with $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ as the response variable. These are three different tip-rate metrics of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one of extinction ($\mu_{BAMM}$, see Methods for details). Predictors included one measure of sexual selection (one of two measures of sexual dichromatism or the index of male-biased sexual selection), four measures of environmental variation, range size and 2-way interactions between the measure of sexual selection and each of the environmental measures including range size, with subsequent model simplification using AIC model selection. In all cases the best model had a $\Delta$ AICc above 20 against the second model, and an evidence ratio above 100.

We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ maximum credibility (MCC) tree ($\beta$ = `r sanitize.numbers(formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b). However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; $\lambda_{BAMM}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b). When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).

@@

Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % HPD interval across PGLS model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC tree. This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}). **Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree. However, our results suggest that BAMM estimates between alternative, similarly plausible phylogenies vary substantially.** Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017). Full details of the BAMM results are presented as supplementary materials.

@@

>>>>

---
title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds
author: 
bibliography: References/SS_Extinction.bib
output:
  pdf_document:
    keep_tex: yes
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
  \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
  \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{}
  \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
  \raggedbottom
csl: References/evolution.csl
nocite: | 
  @Kraaijeveld_2011, @Maia_2013, @Huang_2014, @Gomes_2016, @Cooney_2017, @Janicke_2018, @Mason_2017, @Iglesias-Carrasco_2019, @Hosner_2020, @Price-Waldman_2020
subparagraph: yes
abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates. However, the net effect of sexual selection on diversification is hard to predict, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes. Theoretical work also suggests that the net effect of sexual selection on diversification should depend strongly on ecological factors, though this prediction has seldom been tested. Here, we test whether variation in sexual selection can predict speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera) and whether this relationship is mediated by environmental factors. Male-biased sexual selection, and specifically sexual size dimorphism, predicted two of the three measures of speciation rates that we examined. The link we observed between sexual selection and speciation was independent of environmental variability, though species with smaller ranges had higher speciation rates. There was no association between any proxies of sexual selection and extinction rate. Our findings support the view that male-biased sexual selection, as measured by frequent predictors of male-male competition, has shaped diversification in the largest radiation of birds.
---
<!-- \maketitle\unmarkedfntext{\textsuperscript{†}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia \\ -->
<!-- \textsuperscript{‡}School of Natural and Computational Sciences, Massey University (Albany Campus), Auckland, New Zealand \\ -->
<!-- \textsuperscript{§}\textit{justin.g.cally@gmail.com}} -->

@@

Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data [@Jetz_2012], together with significant advances in phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification. Furthermore, the diversity of outcomes and approaches in previous studies suggests that the association between species diversity and sexual selection is far from clear (reviewed in @tsuji2020sexual).

@@

Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014] because their biology and phylogenetic relationships are comparatively well-known. A 2011 meta-analysis covering 20 primary studies of birds and other taxa found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects or fish [@Kraaijeveld_2011]. However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers. More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019]. **We summarize the major findings from previous studies testing the association between sexual selection and speciation in birds and other taxa since @Kraaijeveld_2011 meta-analysis (Table 1).**

Here, we investigate the association between sexual selection and diversification in birds while building upon previous approaches in multiple ways. We use two measures of the strength of sexual selection: sexual dichromatism [@Dale_2015], as well as an index of male-biased sexual selection [@Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care. We use these two measures because sexual dichromatism does not always signal the presence of strong sexual selection and *vice versa* [@Dale_2015]. For example, male and female dunnocks (*Prunella modularis*) are similarly coloured yet sexual selection appears to be strong [@davies1986re]. Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018]. Additionally, our analysis includes multiple ecological and environmental variables, allowing us to control for potential confounds, **to identify environmental factors, including spatial and temporal environmental variability, interact with sexual selection as theory predicts** [@Connallon_Hall_2016].

@@

**Sexual dichromatism is likely to be imperfectly correlated with variation in the strength of sexual selection across taxa. For this reason, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the ‘index of male-biased sexual selection'.** This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care). The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015]. This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).  \phantomsection

@@

Diversification is the result of two processes, speciation and extinction through time. To estimate speciation rates, we first obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny. Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES; also referred to as diversification rate/DR) is derived from the sum of edge lengths from each tip towards the root, with each edge towards the root having the length down-weighted [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018]. Crucially, studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification. Because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018]. Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path. These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE); but, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].

@@

**We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds; \autoref{Speciation_Figure})**. We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ from the maximum clade credibility (MCC) tree ($\beta$ = `r sanitize.numbers(formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b). However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; $\lambda_{BAMM}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b). When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).

@@

Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % HPD interval across PGLS model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC tree. This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}). **The great majority of earlier studies have based their estimates on a single consensus tree due to the computational requirements of BAMM. However, our results suggest that BAMM estimates between alternative, similarly plausible phylogenies vary substantially.** Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). The calculation of BAMM rates can be affected by the settings of the run and the use of different priors. We therefore compared the estimate of our MCC tree with that of previous published analyses on birds and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017). Full details of the BAMM results are presented as supplementary materials.

@@

<!-- \phantomsection -->
<!-- # Author Contributions   -->

<!-- All authors contributed to conception, design and approach to analyses; J.G.C. and J.D. compiled the data; J.G.C. and I.M. conducted analyses; J.G.C. wrote the first draft of the manuscript; all authors contributed substantially to further manuscript revisions. -->

<!-- \phantomsection -->
<!-- # Data Accessibility -->
<!-- We have sought to make the data and Supplementary Material (including `R` code) freely accessible to readers. The code and Supplementary Material for this manuscript are available on \href{https://justincally.github.io/SexualSelection_Speciation/}{GitHub} and as an attached HTML document (Supplementary Material) for reviewers. Data is archived in the adjoined GitHub repository and will be made publicly available on acceptance.  -->
